---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Import some data

I want to import `FAS_litters.csv`.

```{r}
litters_df = read.csv("data/FAS_litters.csv")
```

Yay! I imported the dataset. Now I want better names.

janitor::clean_names(litters_df) --> using the function without loading the entire package. 
Or load janitor package using library

```{r}
names(litters_df)
litters_df = janitor::clean_names(litters_df)
```

Yay! Now I have better names. Let's look at the dataset. 

## Take a look at the data 

printing in the console. 

```{r}
litters_df

# print out the first few rows of the data.
head(litters_df)

#print out the last few rows of the data.
tail(litters_df)
```

One I use a lot is:

```{r}
view(litters_df) ## put it in the console often not the code chunk, might make knitting process a little difficult.
```

Here's `skimr`: 

```{r}
skimr::skim(litters_df)
```

## Arguments in`read_csv`

```{r,eval=FALSE}
litters_df = 
  read_csv(
    "data/FAS_litters.csv",
    skip = 5,
    col_names = FALSE, 
    show_col_types = FALSE, 
    na = "Low8")
```

check out `?read_csv()` for more information. 

## Parsing columns

Not listing every single column. Specify a single column
```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(`GD of Birth` = col_integer()))
```  

## Reading from Excel 

Reading MLB data.

range --> reading from A1 through F7 subtable.

```{r}
mlb11_df = read_excel("data/mlb11.xlsx", range = "A1:F7")
mlb11_df
```

LorR Words is next.

```{r}
fellow_df = read_excel("data/LotR_Words.xlsx",range = "B3:D6")
```

## Read a SAS file. 

This is where I tried to read a SAS file. Given out first 10 rows.

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
pulse_df
```

