---
title: "Tidy Data"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## pivot longer

Load the PULSE data. 

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Let's try to pivot

```{r}
pulse_tidy = 
  pulse_df %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",  ## getting rid off the name in the column
    values_to = "bdi"
  ) %>% 
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit) ## converting visit to factor variable
  )
```


## pivot_wider

make up a results data table

```{r}
analysis_df = 
  tibble(
    group = c("treatment", "treatment", "control", "control"),
    time = c("a", "b", "a", "b"),
    group_mean = c(4, 8, 3, 6)
  )

analysis_df %>% 
  pivot_wider(
    names_from = "time", 
    values_from = "group_mean"
  ) %>% 
  knitr::kable() ## knitr make it into a table
```


## bind_rows

import the LotR movie words stuff

```{r}
fellowship_df = 
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_rings") 

two_towers_df = 
  read_excel("data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_df = 
  read_excel("data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")

lotr_df = 
  bind_rows(fellowship_df, two_towers_df, return_df) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words"
  ) %>% 
  mutate(race = str_to_lower(race)) %>% 
  relocate(movie)
```

`rbind` is notably worse than `bind_rows`. never use `rbind()`, always use `bind_rows`.


## joins

Look at FAS data. This imports and cleans litters and pups data.

```{r}
# a problem with litters dataset is the group column, there are two variables in the column that are squished together

litters_df = 
  read_csv("data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "day_of_tx"), 3) %>% 
  relocate(litter_number) %>% 
  mutate(dose = str_to_lower(dose))
  
  
pups_df = 
  read_csv("data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    sex = recode(sex, `1` = "male", `2` = "female"), ## putting `` makes r realize it's variable not a number
    sex = factor(sex)) 

```

Let's join these up!

```{r}
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number") %>% 
  relocate(litter_number, dose, day_of_tx)
```

`gather` and `spread` instead of `pivot_longer` and `pivot_wider`. The new functions were updated for good reasons; `gather` and `spread` will still exist, but they’re going to be less common over time.

## Learning Assignments 

This code chunk will import `surv_os` and `surv_program_git` data and cleans both datasets, and then joins them.

```{r}
surv_os = read_csv("survey_results/surv_os.csv") %>% 
  janitor::clean_names() %>% 
  rename(id = what_is_your_uni, os = what_operating_system_do_you_use)

surv_pr_git = read_csv("survey_results/surv_program_git.csv") %>% 
  janitor::clean_names() %>% 
  rename(
    id = what_is_your_uni, 
    prog = what_is_your_degree_program,
    git_exp = which_most_accurately_describes_your_experience_with_git)

left_join(surv_os, surv_pr_git)
inner_join(surv_os, surv_pr_git)
anti_join(surv_os, surv_pr_git)
anti_join(surv_pr_git, surv_os)
```

